---
title: "BMI Relapse Univariate Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggplot2)

#This code reads in the relapse outcomes desired, performs univariate analysis for relapse outcomes, then multivariate analysis of early B-cell repopulation vs. relapse outcomes
#need to have updated "bmi_bcell_relapse_outcomes.csv" first
```

```{r}
database <- read.csv('database.csv')
database
```

```{r}
#This is ALL relapses and the outcomes annualized_relapse, n (total number of annotated relapses), and relapse_time_gap
all_relapses <- read.csv("bmi_bcell_relapse_outcomes.csv")
all_relapses <- all_relapses %>% select(-X, -dob) %>% unique()
all_relapses
```

```{r}
all_repop <- read.csv("all_repop.csv") #reading in repop data frame
all_repop
```

#merging all_repop and all_relapses for analysis 
```{r}
relapse_outcome2 <- merge(all_repop, all_relapses, by = "id_participant", all.x = TRUE)
relapse_outcome2
```
```{r}
relapse_outcome <- relapse_outcome2 %>% 
  select(-X, -age.x) %>% 
  rename(cd19_date = blood_date) #renaming
relapse_outcome$relapse_date <- ymd(relapse_outcome$relapse_date) #reformatting date columns
relapse_outcome$date_msdx <- ymd(relapse_outcome$date_msdx)
relapse_outcome
```

```{r}
#I just use this to check the total number of participants
relapse_outcome %>% drop_na(n) %>% select(id_participant) %>% unique()
```

#making the relapse_time_gap = 2 years for all participants who didn't experience any relapses in the follow-up window
```{r}
relapse_outcome2 <- relapse_outcome %>% 
  drop_na(n)
relapse_outcome2$relapse_time_gap[is.na(relapse_outcome2$relapse_time_gap)] <- 2
outcome_df <- relapse_outcome2
outcome_df
```

#dichot variables, univariate analysis
```{r}
#sex v. n
#res<-t.test(n ~ subject_sex, data=outcome_df)

#sex v. ann relapse rate
res<-t.test(annualized_relapse ~ subject_sex, data=outcome_df)

#sex v. time to relapse
#res<-t.test(relapse_time_gap ~ subject_sex, data=outcome_df)

#race eth v n
#res<-t.test(n ~ race_eth, data=outcome_df)

#race eth v ann relapse rate
#res<-t.test(annualized_relapse ~ race_eth, data=outcome_df)

#race eth v time to relapse
#res<-t.test(relapse_time_gap ~ race_eth, data=outcome_df)
res
```


#cont variables, univariate analysis
```{r}
#age v n
#model = lm(n ~ age, data = outcome_df)

#age v ann relapse rate
#model = lm(annualized_relapse ~ age, data = outcome_df)

#age v time to relapse
#model = lm(relapse_time_gap ~ age, data = outcome_df)

#disease duration v n
#model = lm(n ~ dx_dur, data = outcome_df)

#disease duration v ann relapse rate
#model = lm(annualized_relapse ~ dx_dur, data = outcome_df)

#disease duration v time to relapse
#model = lm(relapse_time_gap ~ dx_dur, data = outcome_df)

#avg diffDate v n
#model = lm(n ~ diffDate_mo, data = outcome_df)

#avg diffDate v ann relapse rate
#model = lm(annualized_relapse ~ diffDate_mo, data = outcome_df)

#avg diffDate v time to relapse
#model = lm(relapse_time_gap ~ diffDate_mo, data = outcome_df)

#baseline cd19 v n
#model = lm(n ~ baseline_cd19_percent, data = outcome_df)

#baseline cd19 v ann relapse rate
model = lm(annualized_relapse ~ baseline_cd19_percent, data = lizzy)

#baseline cd19 v time to relapse
#model = lm(relapse_time_gap ~ baseline_cd19_percent, data = outcome_df)

#baseline bmi v n
#model = lm(n ~ baseline_bmi, data = outcome_df)

#baseline bmi v ann relapse rate
#model = lm(annualized_relapse ~ baseline_bmi, data = outcome_df)

#baseline bmi v time to relapse
#model = lm(relapse_time_gap ~ baseline_bmi, data = outcome_df)
summary(model)
confint(model)
```

```{r}
#making one row per participant, separating by DMT 
bees <- outcome_df %>% 
  select(id_participant, n, annualized_relapse, relapse_time_gap, age, dx_dur, repop_overall, bcell_type, subject_sex) %>% 
  unique() %>% 
  mutate(time_to_relapse_mo = relapse_time_gap/30.42)
bees

ocr_bees <- bees %>% 
  filter(bcell_type == 1)

ritux_bees <- bees %>% 
  filter(bcell_type == 2 | bcell_type == 3 | bcell_type == 4)
```


#actual repop / relapse analysis!
#trying linear regression between repop (0 = no, 1 = yes) & n (relapse number) to start
#adjust covariates as is necessary here
```{r}
#all DMTs combined
model <- lm(n ~ repop_overall + subject_sex + dx_dur, data = bees)
#summary(model)
#confint(model)

#only OCr prts
#model <- lm(n ~ repop_overall + subject_sex + dx_dur, data = ocr_bees)
#summary(model)
#confint(model)

#only rit rux prts
#model <- lm(n ~ repop_overall + subject_sex + dx_dur, data = ritux_bees)
summary(model)
confint(model)

```

```{r}
#I use this to check the number of participants
ritux_bees %>% select(id_participant) %>% unique()
```

#trying annualized relapse rate (calculated by finding the number of years elapsed between date of first neurological symptom to date of CD19 blood draw, then dividing the total number of relapses by that many years)
#using linear regression again
```{r}
#all DMTs combined
#model <- lm(annualized_relapse ~ repop_overall + subject_sex + dx_dur, data = bees)
#summary(model)
#confint(model)

#ocr
#model <- glm(annualized_relapse ~ repop_overall +subject_sex + dx_dur, data = ocr_bees)
#summary(model)
#confint(model)

#rux ritux
model <- glm(annualized_relapse ~ repop_overall + subject_sex + dx_dur, data = ritux_bees)
summary(model)
confint(model)
```

#multivariate analysis, time_to_relapse_mo as outcome
```{r}
#all DMTs combined
#model <- glm(time_to_relapse_mo ~ repop_overall + subject_sex + dx_dur, data = bees)
#summary(model)
#confint(model)

#ocr
#model <- glm(time_to_relapse_mo ~ repop_overall + subject_sex + dx_dur, data = ocr_bees)
#summary(model)
#confint(model)

#rux ritux
model <- glm(time_to_relapse_mo ~ repop_overall + subject_sex + dx_dur, data = ritux_bees)
summary(model)
confint(model)
```


#outcome distribution overview below
```{r}
library(ggplot2)
#outcome_df %>% ggplot(aes(x = n)) +
  #geom_histogram()

#number of participants who have repopulation event
#outcome_df %>% filter(repop == 0) %>% select(id_participant) %>% unique()

#number of participants who have did NOT have repopulation event
#outcome_df %>% filter(repop == 1) %>% select(id_participant) %>% unique()

#number of participants who had at least one relapse
#outcome_df %>% filter(relapse == 1) %>% select(id_participant) %>% unique()

#number of participants who have not had at least one relapse
#outcome_df %>% filter(relapse == 0) %>% select(id_participant) %>% unique()
```

#now trying logistic regression using both binary outcomes for repopulation AND relapse
#note: didn't end up using this in the final analysis/paper
```{r}
#all participants combined
#model <- glm(relapse ~ repop_overall + age + dx_dur + subject_sex + race_eth + bcell_type, data = outcome_df, family = gaussian)
#summary(model)
#confint(model)

#ocr participants
#model <- glm(relapse ~ repop + age + dx_dur + subject_sex + race_eth, data = ocr_outcome, family = gaussian)
#summary(model)
#confint(model)

#rux ritux
model <- glm(relapse ~ repop + age + dx_dur + subject_sex + race_eth + bcell_type, data = rux_ritux_outcome, family = gaussian)
summary(model)
confint(model)
```


#annualized relapse rate vs. overall repopulation at participant level
#again, didn't end up using this in final analysis/paper
```{r}
#all DMTs combined
model <- lm(ann_relapse_rate ~ repop_overall + age + dx_dur + subject_sex + race_eth, data = outcome_df)
summary(model)
#confint(model)

#ocr
#model <- glm(ann_relapse_rate ~ repop_overall + age, data = ocr_outcome)
#summary(model)
#confint(model)

#rux ritux
#model <- glm(ann_relapse_rate ~ repop_overall + bcell_type + age, data = rux_ritux_outcome)
#summary(model)
#confint(model)
```




















