---
title: "BMI B-cell PDDS Sustained Disability Outcomes"
output: html_document
date: "2023-09-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
database <- read.csv('database.csv')
database
```


#now reading in confirmed disability code output
```{r}
#Here make sure to read in the right output code depending on whether you're running confirmed or sustained disability progression
##If running confirmed disability progression, line below is: df_sustain <- read.csv("conf_disability_output.csv")
##If running sustained disability progression, line below is: df_sustain <- read.csv("df_sustain.csv")
df_sustain <- read.csv("conf_disability_output.csv")
df_sustain <- df_sustain %>% 
  select(-X)
df_sustain
#data key!
## 0 = had sufficient data to assess, neither sustained increase or decrease
## 1 = had sufficient data to assess, sustained PDDS increase (disability worsening)
## 2 = had sufficient data to assess, sustained PDDS decrease (disability improvement)
## NA = did not have sufficient PDDS data to assess
```

```{r}
yn_repop2 <- read.csv("all_repop.csv") #reading in repop data
yn_repop <- yn_repop2 %>% 
  rename(first_infusion_date = baseline_infusion_date) %>% #renaming
  rename(treat_dur = treatment_dur_mo)
yn_repop
```



#making binary outcomes for both repopulation (any_repop_occur == 1 if the participant had at least one repopulation event, sustain_worsening == 1 if the participant had a confirmed period of sustained disability worsening [pdds increase], etc.)
```{r}
pdds_repop2 <- merge(yn_repop, df_sustain, by = "id_participant")
pdds_repop <- pdds_repop2 %>% 
  group_by(id_participant) %>% 
  #mutate(any_repop_occur = ifelse(any(repop == 1), 1, 0)) %>% 
  ungroup() %>% 
  drop_na(disability_status)
pdds_repop$first_infusion_date <- ymd(pdds_repop$first_infusion_date)
pdds_repop <- pdds_repop %>% 
  group_by(id_participant) %>% 
  mutate(sustain_worsening = ifelse(any(disability_status == 1), 1, 0)) %>% 
  mutate(sustain_improvement = ifelse(any(disability_status == 2), 1, 0)) %>% 
  ungroup() %>% 
  #select(-X) %>% 
  unique()
pdds_repop$baseline_blood_date <- ymd(pdds_repop$baseline_blood_date)
baseline_actual <- pdds_repop %>% 
  select(id_participant, baseline_blood_date, baseline_cd19_percent, first_infusion_date) %>% 
  filter(baseline_blood_date < first_infusion_date) %>% 
  select(-first_infusion_date) %>% 
  unique()
baseline_actual
pdds_repop <- pdds_repop %>% 
  select(-baseline_blood_date, -baseline_cd19_percent)
pdds_repop3 <- merge(pdds_repop, baseline_actual, by = "id_participant", all.x = TRUE)
pdds_repop <- pdds_repop3
pdds_repop 
```


#starting univariate analysis (continuous vars), separating by DMT
```{r}
outcome_df <- pdds_repop %>% drop_na(disability_status) %>% unique()
ocr_outcome <- outcome_df %>% dplyr::filter(bcell_type == 1)%>% unique()
rux_ritux_outcome <- outcome_df %>% dplyr::filter(bcell_type == 2 | bcell_type == 3 | bcell_type == 4)%>% unique()
outcome_df
```

```{r}
#I use this chunk to check number of participants
rux_ritux_outcome %>% select(id_participant) %>% unique()
```

#dichot variables (Fisher exact test to compare categorical v categorical/binary)
```{r}
#sex v. disability worsening
#looky <- outcome_df %>% select(id_participant, subject_sex, sustain_worsening) %>% unique()
#contingency_table <- table(looky$sustain_worsening, looky$subject_sex)

#sex v. disability improvement
#looky <- outcome_df %>% select(id_participant, subject_sex, sustain_improvement) %>% unique()
#contingency_table <- table(looky$sustain_improvement, looky$subject_sex)

#race eth v disability improvement
#looky <- outcome_df %>% select(id_participant, race_eth, sustain_improvement) %>% unique()
#contingency_table <- table(looky$sustain_improvement, looky$race_eth)

#race eth v disability worsening
looky <- outcome_df %>% select(id_participant, race_eth, sustain_worsening) %>% unique()
contingency_table <- table(looky$sustain_worsening, looky$race_eth)

#repop_overall v. disability worsening
#looky <- outcome_df %>% select(id_participant, repop_overall, sustain_worsening) %>% unique()
#contingency_table <- table(looky$sustain_worsening, looky$repop_overall)

#repop_overall v. disability improve
#looky <- outcome_df %>% select(id_participant, repop_overall, sustain_improvement) %>% unique()
#contingency_table <- table(looky$sustain_improvement, looky$repop_overall)

#repop v. disability worsening
#looky <- outcome_df %>% select(id_participant, repop, sustain_worsening) %>% unique()
#contingency_table <- table(looky$sustain_worsening, looky$repop)

#repop v. disability improvement
#looky <- outcome_df %>% select(id_participant, repop, sustain_improvement) %>% unique()
#contingency_table <- table(looky$sustain_improvement, looky$repop)

res<-fisher.test(contingency_table)
res
contingency_table
```

```{r}
#generating a contigency table to see data distribution
looky <- outcome_df %>% 
  select(id_participant, repop_overall, sustain_worsening) %>% 
  unique()
looky
contingency_table <- table(looky$sustain_worsening, looky$repop_overall)
contingency_table
```


#univariate analysis, continuous variables v binary outcome using logistic regression
```{r}
#age v. sustained worsening
#model <- glm(sustain_worsening ~ age, data = outcome_df, family = binomial)

#age v. sustained improvement
#model <- glm(sustain_improvement ~ age, data = outcome_df, family = binomial)

#dx_dur v. sustained worsening
#model <- glm(sustain_worsening ~ dx_dur, data = outcome_df, family = binomial)

#dx_dur v. sustained improvement
#model <- glm(sustain_improvement ~ dx_dur, data = outcome_df, family = binomial)

#avg_diffDate v. sustained worsening
#model <- glm(sustain_worsening ~ diffDate_mo, data = outcome_df, family = binomial)

#avg_diffDate v. sustained improvement
#model <- glm(sustain_improvement ~ diffDate_mo, data = outcome_df, family = binomial)

#baseline BMI v. sustained worsening
#model <- glm(sustain_worsening ~ baseline_bmi, data = outcome_df, family = binomial)

#baseline BMI v. sustained improvement
#model <- glm(sustain_improvement ~ baseline_bmi, data = outcome_df, family = binomial)

#baseline CD19% v. sustained worsening
model <- glm(sustain_worsening ~ baseline_cd19_percent, data = lizzy, family = binomial)

#baseline CD19% v. sustained improvement
#model <- glm(sustain_improvement ~ baseline_cd19_percent, data = outcome_df, family = binomial)
summary(model)
confint(model)
```


```{r}
#narrowing down to one row per participant, adjust necessary covars here, separating by DMT
outcome_df15 <- outcome_df %>% 
  select(id_participant, sustain_worsening, sustain_improvement, repop_overall, bcell_type, age) %>% 
  unique()
ocr_outcome5 <- outcome_df15 %>% dplyr::filter(bcell_type == 1)%>% unique()
rux_ritux_outcome5 <- outcome_df15 %>% dplyr::filter(bcell_type == 2 | bcell_type == 3 | bcell_type == 4)%>% select(-bcell_type)%>% unique()
outcome_df5 <- outcome_df15 %>% 
  select(-bcell_type) %>% unique()
outcome_df5
```

#final analysis: Logistic regression Sustained / confirmed worsening v. any_repop_occur including significant covariates
```{r}
#sustained worsening v. any_repop_occur ALL PRTs
model <- glm(sustain_worsening ~ repop_overall + age, data = outcome_df5, family = binomial)
#summary(model)
#confint(model)

#sustained worsening v. any_repop_occur OCR PRTs
#model <- glm(sustain_worsening ~ repop_overall + age, data = ocr_outcome5, family = binomial)
#summary(model)
#confint(model)

#sustained worsening v. any_repop_occur ritux PRTs
#model <- glm(sustain_worsening ~ repop_overall + age, data = rux_ritux_outcome5, family = binomial)
summary(model)
confint(model)
```

```{r}
#just use this to check number of participants
outcome_df5  %>% select(id_participant) %>% unique()
```


```{r}
#this is same as outcome_df5, but uses repop at an event level instead of at a participant level (e.g. repop instead of repop_overall)
outcome_df6 <- outcome_df %>% 
  select(id_participant, sustain_worsening, sustain_improvement, repop, bcell_type, age, baseline_bmi, dx_dur) %>% 
  unique()
outcome_df6
ocr_outcome6 <- outcome_df6 %>% dplyr::filter(bcell_type == 1)%>% unique()
rux_ritux_outcome6 <- outcome_df6 %>% dplyr::filter(bcell_type == 2 | bcell_type == 3 | bcell_type == 4)%>% unique()
```



#final analysis: Logistic regression Sustained worsening v. repop including significant covariates (event level)
```{r}
#sustained worsening v. repop ALL PRTs
model <- glm(sustain_worsening ~ repop + age + baseline_bmi, data = outcome_df6, family = binomial)
#summary(model)
#confint(model)

#sustained worsening v. repop OCR PRTs
#model <- glm(sustain_worsening ~ repop + age + baseline_bmi, data = ocr_outcome6, family = binomial)
#summary(model)
#confint(model)

#sustained worsening v. repop OCR PRTs
#model <- glm(sustain_worsening ~ repop + age + baseline_bmi, data = rux_ritux_outcome6, family = binomial)
summary(model)
confint(model)
```


```{r}
rux_ritux_outcome %>% select(id_participant) %>% unique()
```
#final analysis: Logistic regression Sustained Improvement v. any_repop_occur including significant covariates
```{r}
#sustained worsening v. repop_overall ALL PRTs
#model <- glm(sustain_improvement ~ repop_overall + age + dx_dur + baseline_bmi, data = outcome_df5, family = binomial)
#summary(model)
#confint(model)

#sustained worsening v. repop_overall OCR PRTs
#model <- glm(sustain_improvement ~ repop_overall + age + dx_dur + baseline_bmi, data = ocr_outcome5, family = binomial)
#summary(model)
#confint(model)

#sustained worsening v. repop_overall OCR PRTs
model <- glm(sustain_improvement ~ repop_overall + age + dx_dur + baseline_bmi, data = rux_ritux_outcome5, family = binomial)
summary(model)
confint(model)
```

#final analysis: Logistic regression Sustained Improvement v. repop including significant covariates (event level)
```{r}
#sustained worsening v. any_repop_occur ALL PRTs
#model <- glm(sustain_improvement ~ repop , data = outcome_df6, family = binomial)
#summary(model)
#confint(model)

#sustained worsening v. any_repop_occur OCR PRTs
#model <- glm(sustain_improvement ~ repop , data = ocr_outcome6, family = binomial)
#summary(model)
#confint(model)

#sustained worsening v. any_repop_occur OCR PRTs
model <- glm(sustain_improvement ~ repop , data = rux_ritux_outcome6, family = binomial)
summary(model)
confint(model)
```



#final analysis: Logistic regression Sustained Improvement v. any_repop_occur including significant covariates (adjusted for baseline bmi only)
#note: didn't end up using this in final paper
```{r}
#sustained worsening v. any_repop_occur ALL PRTs
#model <- glm(sustain_improvement ~ any_repop_occur + baseline_bmi, data = outcome_df, family = binomial)
#summary(model)
#confint(model)

#sustained worsening v. any_repop_occur OCR PRTs
#model <- glm(sustain_improvement ~ any_repop_occur + baseline_bmi, data = ocr_outcome, family = binomial)
#summary(model)
#confint(model)

#sustained worsening v. any_repop_occur OCR PRTs
model <- glm(sustain_improvement ~ any_repop_occur + baseline_bmi, data = rux_ritux_outcome, family = binomial)
summary(model)
confint(model)
```








